---
title: "Agriculture Sector Analysis using CRISP-DM Framework"
author: "Wong Yi Ting (S2152880)"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
runtime: shiny
---

# **The Role of Data Science in Agriculture Sector**

## **Introduction**
Data science enables better understanding and management of agricultural price volatility by analyzing trends, predicting fluctuations, and providing actionable insights. This study specifically focuses on leveraging predictive models to assist policymakers and farmers in making informed decisions to optimize production strategies and mitigate risks.

---

## **CRISP-DM Framework Analysis**

### **1. Business Understanding**
Volatile agricultural prices directly impact farmers' income and decision-making processes. This study aims to:<br>
&emsp;**1.** Predict future agriculture PPI values using time series models.<br>
&emsp;**2.** Use other industries' PPI data to predict agriculture PPI using regression models.<br>

These objectives address critical industry needs, such as stabilizing market conditions and improving income predictability for stakeholders.

### **2. Data Understanding**
- **Dataset Source**: Department of Statistics Malaysia
- **Fields**: 
  - **`date`**: Monthly timestamps
  - **`agriculture`**: Agriculture PPI
  - **`mining`**: Mining PPI
  - **`manufacturing`**: Manufacturing PPI
  - **`electricity`**: Electricity & Gas Supply PPI
  - **`water`**: Water Supply PPI

```{r data-summary, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)

# Load dataset
data <- read.csv("ppi.csv")

# Summary of the dataset
summary(data)

# Check for missing values
cat("Missing values in dataset:", sum(is.na(data)), "\n")

# Ensure agriculture data is numeric and handle missing values
data$agriculture <- as.numeric(as.character(data$agriculture))
data <- data %>% filter(!is.na(agriculture))
cat("Total rows after cleaning:", nrow(data), "\n")
```

#### **Data Visualizations**

```{r data-visualizations, message=FALSE, warning=FALSE}
# Plot Agriculture PPI over time
ggplot(data, aes(x = as.Date(date), y = agriculture)) +
  geom_line(color = "#2E8B57", size = 1.2) +  
  ggtitle("Agriculture PPI Over Time from 2010 to 2024") +
  xlab("Date") +
  ylab("Agriculture PPI") +
  theme_minimal(base_size = 14) +            
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), 
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    panel.grid.major = element_line(color = "gray80", linetype = "dashed")
  )

# Plot correlation between Agriculture PPI and Mining PPI
correlation_matrix <- cor(
  data[, c("agriculture", "mining", "manufacturing", "electricity", "water")], 
  use = "complete.obs"
)
correlation_data <- melt(correlation_matrix)
ggplot(correlation_data, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +  # Add white borders for tiles
  scale_fill_gradient2(
    low = "#2b83ba", high = "#d7191c", mid = "#f7f7f7", 
    midpoint = 0, limit = c(-1, 1), name = "Correlation"
  ) +  # Improved color scheme
  ggtitle("Heatmap of Correlations with Agriculture PPI") +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )


# Plot distribution of Agriculture PPI
ggplot(data, aes(x = agriculture)) +
  geom_histogram(binwidth = 5, fill = "#FF8C00", color = "white", alpha = 0.9) +  # Soft orange with white borders
  ggtitle("Distribution of Agriculture PPI") +
  xlab("Agriculture PPI") +
  ylab("Frequency") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),  # Centered and bold title
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    panel.grid.major = element_line(color = "gray80", linetype = "dashed")
  )
```

### **3. Data Preparation**
Prepare the dataset for both objectives.

#### **3.1 Time Series Data Preparation**
Prepare the agriculture PPI data for time series analysis.

```{r ts-preparation, message=FALSE, warning=FALSE}
# Create a time series object for agriculture PPI
agriculture_ts <- ts(data$agriculture, start = c(2010, 1), frequency = 12)
```

#### **3.2 Regression Data Preparation**
Prepare the dataset to predict agriculture PPI based on other industries' PPI data.

```{r regression-preparation, message=FALSE, warning=FALSE}
# Select relevant columns for regression
data_regression <- data %>%
  select(agriculture, mining, manufacturing, electricity, water) %>%
  na.omit()

# Split data into training and testing sets
set.seed(123)
library(caret)
trainIndex <- createDataPartition(data_regression$agriculture, p = 0.8, list = FALSE)
trainData <- data_regression[trainIndex, ]
testData <- data_regression[-trainIndex, ]
```

#### **3.3 Feature Selection**
Analyze feature importance using Random Forest and select important features.

```{r feature-selection, message=FALSE, warning=FALSE}
library(randomForest)

# Train Random Forest model to calculate feature importance
rf_model_importance <- randomForest(
  agriculture ~ ., 
  data = trainData, 
  ntree = 500, 
  mtry = 2, 
  importance = TRUE
)

# Print feature importance
importance <- randomForest::importance(rf_model_importance)
print(importance)

# Visualize feature importance
varImpPlot(rf_model_importance)

# Select important features based on importance scores
selected_features <- c("agriculture", "mining", "manufacturing", "electricity")

# Update training and testing datasets with selected features
trainData_selected <- trainData[, selected_features]
testData_selected <- testData[, selected_features]
```

### **4. Modeling**

#### **4.1 Time Series Modeling: Multi-Model**
Encapsulate ARIMA and ETS models into reusable functions.

```{r ts-function, message=FALSE, warning=FALSE}
library(forecast)

# Function to train ARIMA model
ARIMA_Model <- function(ts_data) {
  model <- auto.arima(ts_data)
  forecast_data <- forecast(model, h = 12)
  rmse <- sqrt(mean((ts_data - fitted(model))^2))
  mae <- mean(abs(ts_data - fitted(model)))
  mape <- mean(abs((ts_data - fitted(model)) / ts_data)) * 100
  r_squared <- 1 - sum((ts_data - fitted(model))^2) / sum((ts_data - mean(ts_data))^2)
  return(list(model = model, forecast = forecast_data, RMSE = rmse,MAE = mae, MAPE = mape, R2 = r_squared))
}

# Function to train ETS model
ETS_Model <- function(ts_data) {
  model <- ets(ts_data)
  forecast_data <- forecast(model, h = 12)
  rmse <- sqrt(mean((ts_data - fitted(model))^2))
  mae <- mean(abs(ts_data - fitted(model)))
  mape <- mean(abs((ts_data - fitted(model)) / ts_data)) * 100
  r_squared <- 1 - sum((ts_data - fitted(model))^2) / sum((ts_data - mean(ts_data))^2)
  return(list(model = model, forecast = forecast_data, RMSE = rmse,MAE = mae, MAPE = mape, R2 = r_squared))
}

# Apply ARIMA and ETS models to the time series
arima_results <- ARIMA_Model(agriculture_ts)
ets_results <- ETS_Model(agriculture_ts)

cat("ARIMA RMSE:", arima_results$RMSE,"\n","ARIMA MAE:",arima_results$MAE,"\n","ARIMA MAPE:", arima_results$MAPE,"%\n","ARIMA R²:",arima_results$R2,"\n")


cat("ETS RMSE:", ets_results$RMSE,"\n","ETS MAE:",ets_results$MAE,"\n","ETS MAPE:", ets_results$MAPE,"%\n","ETS R²:",ets_results$R2,"\n")
```

#### **4.2 Regression Modeling: Multi-Model**
Encapsulate Linear Regression, Random Forest, and XGBoost into reusable functions.

```{r regression-functions, message=FALSE, warning=FALSE}
library(randomForest)
library(caret)
library(xgboost)
library(class)

# Function to train Linear Regression model
Linear_Regression <- function(train_data, test_data) {
  model <- lm(agriculture ~ ., data = train_data)
  predictions <- predict(model, test_data)
  actual <- test_data$agriculture
  
  rmse <- sqrt(mean((actual - predictions)^2))
  mae <- mean(abs(actual - predictions))
  r2 <- 1 - sum((actual - predictions)^2) / sum((actual - mean(actual))^2)
  mape <- mean(abs((actual - predictions) / actual)) * 100

  return(list(model = model, predictions = predictions, RMSE = rmse, MAE = mae, R2 = r2, MAPE = mape))
}

# Function to train Random Forest model
RF_Model <- function(train_data, test_data, ntree = 1000, mtry = NULL, nodesize = 5) {
  if (is.null(mtry)) {
    mtry <- floor(sqrt(ncol(train_data) - 1))
  }
  model <- randomForest(
    agriculture ~ ., 
    data = train_data, 
    ntree = ntree, 
    mtry = mtry, 
    nodesize = nodesize
  )
  predictions <- predict(model, test_data)
  actual <- test_data$agriculture

  rmse <- sqrt(mean((actual - predictions)^2))
  mae <- mean(abs(actual - predictions))
  r2 <- 1 - sum((actual - predictions)^2) / sum((actual - mean(actual))^2)
  mape <- mean(abs((actual - predictions) / actual)) * 100

  return(list(model = model, predictions = predictions, RMSE = rmse, MAE = mae, R2 = r2, MAPE = mape))
}

# Function to train XGBoost model
XGBoost_Model <- function(train_data, test_data, nrounds = 1000, max_depth = 6, eta = 0.3) {
  # Prepare data for XGBoost
  dtrain <- xgb.DMatrix(data = as.matrix(train_data[, -1]), label = train_data$agriculture)
  dtest <- xgb.DMatrix(data = as.matrix(test_data[, -1]), label = test_data$agriculture)

  # Set XGBoost parameters
  params <- list(
    objective = "reg:squarederror",
    max_depth = max_depth,
    eta = eta
  )

  # Train the XGBoost model
  model <- xgb.train(params = params, data = dtrain, nrounds = nrounds)

  # Predict and calculate metrics
  predictions <- predict(model, dtest)
  actual <- test_data$agriculture

  rmse <- sqrt(mean((actual - predictions)^2))
  mae <- mean(abs(actual - predictions))
  r2 <- 1 - sum((actual - predictions)^2) / sum((actual - mean(actual))^2)
  mape <- mean(abs((actual - predictions) / actual)) * 100

  return(list(model = model, predictions = predictions, RMSE = rmse, MAE = mae, R2 = r2, MAPE = mape))
}

# Function to train KNN Regression model
KNN_Model <- function(data_train, data_test, k = 5) {
  model <- train(agriculture ~ ., data = data_train, method = "knn", tuneGrid = expand.grid(k = k))
  x_test <- data_test[, -1]
  y_test <- data_test$agriculture
  predictions <- predict(model, x_test)

  rmse <- sqrt(mean((y_test - predictions)^2))
  mae <- mean(abs(y_test - predictions))
  r2 <- 1 - sum((y_test - predictions)^2) / sum((y_test - mean(y_test))^2)
  mape <- mean(abs((y_test - predictions) / y_test)) * 100

  return(list(model = model, RMSE = rmse, MAE = mae, R2 = r2, MAPE = mape))
}

# Apply Linear Regression, Random Forest, and XGBoost models
linear_results <- Linear_Regression(trainData_selected, testData_selected)
rf_results <- RF_Model(trainData_selected, testData_selected)
xgboost_results <- XGBoost_Model(trainData_selected, testData_selected, nrounds = 2000, max_depth = 4, eta = 0.01)
knn_results <- KNN_Model(trainData_selected, testData_selected, k = 10)

# Print metrics for all models
cat("Linear Regression Metrics:\n","RMSE:", linear_results$RMSE, "\n","MAE:", linear_results$MAE, "\n","R2:", linear_results$R2, "\n","MAPE:", linear_results$MAPE, "%\n")

cat("Random Forest Metrics:\n","RMSE:", rf_results$RMSE, "\n","MAE:", rf_results$MAE, "\n","R2:", rf_results$R2, "\n","MAPE:", rf_results$MAPE, "%\n")

cat("XGBoost Metrics:\n","RMSE:", xgboost_results$RMSE, "\n","MAE:", xgboost_results$MAE, "\n","R2:", xgboost_results$R2, "\n","MAPE:", xgboost_results$MAPE, "%\n")

cat("KNN Metrics:\n","RMSE:", knn_results$RMSE, "\n","MAE:", knn_results$MAE, "\n","R2:", knn_results$R2, "\n","MAPE:", knn_results$MAPE, "%\n")

```

### **5. Evaluation**

#### **5.1 Time Series Model Selection**
Evaluate and display results for each time series model, then select the best model.

```{r evaluation-timeseries, message=FALSE, warning=FALSE}
if (arima_results$RMSE < ets_results$RMSE) {
  best_ts_model <- "ARIMA"
  best_ts_forecast <- arima_results$forecast
} else {
  best_ts_model <- "ETS"
  best_ts_forecast <- ets_results$forecast
}

cat("Best Time Series Model:", best_ts_model, "\n")
```

#### **5.2 Regression Model Selection**
Evaluate and display results for each regression model, then select the best model.

```{r evaluation-regression, message=FALSE, warning=FALSE}
### Regression Model Selection
if (linear_results$RMSE < rf_results$RMSE & linear_results$RMSE < xgboost_results$RMSE & linear_results$RMSE < knn_results$RMSE) {
  best_reg_model <- "Linear Regression"
  best_reg_predictions <- linear_results$predictions
} else if (rf_results$RMSE < xgboost_results$RMSE & rf_results$RMSE < knn_results$RMSE) {
  best_reg_model <- "Random Forest"
  best_reg_predictions <- rf_results$predictions
} else if (xgboost_results$RMSE < knn_results$RMSE) {
  best_reg_model <- "XGBoost"
  best_reg_predictions <- xgboost_results$predictions
} else {
  best_reg_model <- "KNN"
  best_reg_predictions <- knn_results$predictions
}

cat("Best Regression Model:", best_reg_model, "\n")
```

```{r save-data, include=FALSE}
save(agriculture_ts, testData_selected, rf_results, file = "agriculture_analysis_data.RData")
```

### **6. Deployment**
Visualize the insights through trend graphs, regression results, and innovative predictions.

```{r dynamic-predict, message=FALSE, warning=FALSE}
library(shiny)
library(forecast)
library(ggplot2)
library(zoo)

# Define Shiny app UI
ui <- fluidPage(
  tags$head(
    tags$style(HTML(
      "body {font-family: Arial, sans-serif; background-color: #f8f9fa;}
      .title {color: #000000; text-align: center; margin-bottom: 20px;}
      .sidebar {background-color: #ffffff; padding: 20px; border-radius: 10px; box-shadow: 0px 4px 6px rgba(0,0,0,0.1);}
      .btn-primary {background-color: #007bff; border-color: #007bff;}
      .main-panel {padding: 20px; background-color: #ffffff; border-radius: 10px; box-shadow: 0px 4px 6px rgba(0,0,0,0.1);}"
    ))
  ),
  titlePanel(tags$h1("Dynamic Prediction of Agriculture PPI", class = "title")),
  sidebarLayout(
    sidebarPanel(
      class = "sidebar",
      numericInput("year", "Enter Target Year:", value = 2025, min = 2023, max = 2035),
      numericInput("month", "Enter Target Month:", value = 12, min = 1, max = 12),
      numericInput("mining_adjust", "Adjust Mining PPI (%):", value = 10, min = 0, max = 100, step = 1),
      numericInput("history_points", "Number of Historical Points to Include:", value = 80, min = 50, max = 200),
      numericInput("threshold", "Set Threshold for Volatility Alert:", value = 180, min = 50, max = 200),
      actionButton("predict", "Generate", class = "btn btn-primary"),
      p("Choose the target year, month, and parameters, then click 'Generate'.", style = "color: #6c757d; font-size: 14px;")
    ),
    mainPanel(
      class = "main-panel",
      tabsetPanel(
        tabPanel("Forecast Visualization", plotOutput("forecastPlot", height = "500px")),
        tabPanel("Additional Analysis",
          fluidRow(
            column(12, plotOutput("uncertaintyPlot", height = "400px")),
            column(12, plotOutput("sensitivityPlot", height = "400px"))
          ),
          verbatimTextOutput("alertMessage")
        )
      )
    )
  )
)

# Define Shiny app server logic
server <- function(input, output) {
  forecast_result <- eventReactive(input$predict, {
    arima_model <- auto.arima(agriculture_ts)
    current_date <- Sys.Date()
    target_date <- as.Date(paste0(input$year, "-", input$month, "-01"))
    months_ahead <- as.integer(12 * (as.numeric(format(target_date, "%Y")) - as.numeric(format(current_date, "%Y"))) +
                                  (as.numeric(format(target_date, "%m")) - as.numeric(format(current_date, "%m"))))
    forecast_result <- forecast(arima_model, h = months_ahead)
    return(forecast_result)
  })

  sensitivity_analysis <- reactive({
    sensitivity_test <- testData_selected
    sensitivity_test$mining <- sensitivity_test$mining * (1 + input$mining_adjust / 100)
    sensitivity_predictions <- predict(rf_results$model, sensitivity_test)
    data.frame(Actual = testData_selected$agriculture, Predicted = sensitivity_predictions)
  })

  output$forecastPlot <- renderPlot({
    req(forecast_result())
    autoplot(forecast_result(), series = "Forecast", size = 1) +
      geom_line(color = "#2E8B57", size = 1.2) +
      ggtitle(paste("Dynamic Forecast for Agriculture PPI in", "Year:",input$year, "& Month:",input$month)) +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray80", linetype = "dashed"),
        panel.grid.minor = element_blank()
      ) +
      xlab("Year") + ylab("Agriculture PPI")
  })

  output$uncertaintyPlot <- renderPlot({
    req(forecast_result())
    autoplot(forecast_result(), include = input$history_points) +
      ggtitle("Prediction with Confidence Intervals") +
      xlab("Year") + ylab("Agriculture PPI") +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(face = "bold", size = 16,hjust = 0.5),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray80", linetype = "dashed"),
        panel.grid.minor = element_blank()
      )
  })
  cat("\n")
  output$sensitivityPlot <- renderPlot({
    sensitivity_data <- sensitivity_analysis()
    ggplot(sensitivity_data, aes(x = Actual, y = Predicted)) +
      geom_point() +
      geom_abline(intercept = 0, slope = 1, color = "red") +
      ggtitle("Sensitivity Analysis: Mining PPI Adjusted") +
      xlab("Actual Agriculture PPI") + ylab("Predicted Agriculture PPI") +
      theme_minimal(base_size = 14) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        panel.grid.major = element_line(color = "gray80", linetype = "dashed"),
        panel.grid.minor = element_blank()
      )
  })

  output$alertMessage <- renderPrint({
    req(forecast_result())
    threshold <- input$threshold
    if (any(forecast_result()$mean > threshold)) {
      cat("Alert: Projected PPI exceeds threshold! Consider policy interventions.\n")
    } else {
      cat("No significant price volatility detected.\n")
    }
  })
}
shinyApp(ui = ui, server = server)
```

---

## **Conclusion**
By applying the CRISP-DM framework:<br>
&emsp;- Evaluated ARIMA and ETS models for time series forecasting and selected the best-performing model.<br>
&emsp;- Evaluated Linear Regression, Random Forest, XGBoost models and K-Nearest Neighbors for predicting agriculture PPI and selected the best-performing model.<br>
&emsp;- Conducted sensitivity and uncertainty analyses to provide deeper insights into agricultural price trends.<br>
&emsp;- Developed an alert system to flag significant price volatility, aiding decision-makers.

These models and analyses offer actionable insights for stabilizing agricultural markets, optimizing production, and supporting policy formulation.

---

## **References**
- Dataset: [PPI Data CSV](https://storage.dosm.gov.my/ppi/ppi.csv)
- Packages: `dplyr`, `ggplot2`, `caret`, `forecast`, `randomForest`, `xgboost`, `class`